---
title: "What happens if we only use positional or geometric features?"
author: "Mark Platts"
date: "5th September 2017"
output:
  html_document: default
  html_notebook: default
---

Purpose:
In this analysis we will apply the k-means algorithm twice each time for two subsets of features. In the first we will use only the geometric and in the second we will only look at positional.

Load necessary libraries & sources
```{r}
library(data.table)
library(stringr)
library(ggplot2)
library(here)
source(here("src/R/helper_functions.R"))
```

Load data into memory and check that it doesn't contain any NA's
```{r}
dt <- fread(here("Data/Output_features/segment_features.csv"),
            stringsAsFactors = TRUE)

print(any(is.na(dt)))

dt_xy <- fread(here("Data/Output_features/segment_xys.csv"), head = TRUE)
```
Splitting into datasets with only positional and geometric features
```{r}
descripter_columns   <- c("SegmentID", "Experiment", "UsingLight", "FileName")
geometric_features   <- c("Focus", "Eccentricity", "PathEfficiency", "SumAbsoluteAngles")
positional_features  <- c("MedianDistanceFromCentre", "IQRange", "InnerRadiusVariation", "CentralDisplacement")
```

# Geometric features

Three clusters:
```{r, fig.width=5,fig.height=10}
source(here("src/R/helper_functions.R"))
set.seed(1) #the following code take a random sample of the segments
dt_geometric <- dt[, c(descripter_columns, geometric_features), with = FALSE]
km.out <- kmeans(x = scale(dt_geometric[, (geometric_features), with = FALSE]), centers = 3, nstart = 20)
dt_geometric[, cluster := km.out$cluster]
plot_cluster(features = dt_geometric, xy = dt_xy, nSegments = 5)
```

5 clusters:
```{r, fig.width=9,fig.height=20}
source(here("src/R/helper_functions.R"))
set.seed(1) #the following code take a random sample of the segments
dt_geometric <- dt[, c(descripter_columns, geometric_features), with = FALSE]
km.out <- kmeans(x = scale(dt_geometric[, (geometric_features), with = FALSE]), centers = 5, nstart = 20)
dt_geometric[, cluster := km.out$cluster]
plot_cluster(features = dt_geometric, xy = dt_xy, nSegments = 10)
```

# Positional features

Three clusters:
```{r, fig.width=5,fig.height=10}
source(here("src/R/helper_functions.R"))
set.seed(1) #the following code take a random sample of the segments
dt_positional <- dt[, c(descripter_columns, positional_features), with = FALSE]
km.out <- kmeans(x = scale(dt_positional[, (positional_features), with = FALSE]), centers = 3, nstart = 20)
dt_positional[, cluster := km.out$cluster]
plot_cluster(features = dt_positional, xy = dt_xy, nSegments = 5)
```

Five clusters:
```{r, fig.width=9,fig.height=10}
source(here("src/R/helper_functions.R"))
set.seed(1) #the following code take a random sample of the segments
dt_positional <- dt[, c(descripter_columns, positional_features), with = FALSE]
km.out <- kmeans(x = scale(dt_positional[, (positional_features), with = FALSE]), centers = 5, nstart = 20)
dt_positional[, cluster := km.out$cluster]
plot_cluster(features = dt_positional, xy = dt_xy, nSegments = 5)
```