---
title: "Using the silhouette technique to select the optimum number of clusters and principle components"
author: "Mark Platts"
date: "21st October 2017"
output: html_notebook
---

Load necessary libraries & sources
```{r}
library(data.table)
library(stringr)
library(ggplot2)
library(cluster)
library(here)

source(here("src/R/helper_functions.R"))
```

Load data into memory and check that it doesn't contain any NA's
```{r, message = F}
dt <- fread(here("Data/Output_features/segment_features.csv"),
            stringsAsFactors = TRUE)

print(any(is.na(dt)))

dt_xy <- fread(here("Data/Output_features/segment_xys.csv"), head = TRUE)
```

Since I believe it unlikely that speed will determine the appearance of a segment for now we will exclude it.

```{r}
dt_names <- names(dt)
speed_names <- str_subset(dt_names, "Speed")
dt_mod <- dt[, (speed_names) := NULL]
```

There is likely a lot of correlation of variables. To improve efficiency of calculating the kmeans I will use PCA

```{r}
pr.out <- prcomp(dt_mod[, MedianDistanceFromCentre:SumAbsoluteAngles], scale = TRUE)
```

Looks like the first 2 components will be pretty good. lets take a look at what it looks like in 2-d plot.

```{r}
plot(pr.out$x[, "PC1"], pr.out$x[, "PC2"], pch = ".")
```

Let's zoom in and see if we can see anything interesting happening:
```{r}
plot(pr.out$x[, "PC1"], pr.out$x[, "PC2"], pch = ".", xlim = c(-5,7))
```

```{r}
plot(pr.out$x[, "PC1"], pr.out$x[, "PC2"], pch = ".", xlim = c(-5,0), ylim = c(-3,1))
```

There isn't any clear clustering going on here, but something unusual is happening. I do think that we will get a clearer picture once a voting system has been applied to each point to determine the cluster it is in and then to plot them; this should clear away some of the points inbetween clusters and create a clearer distinction.

Let's try the silhouette technique to select number of principal components and clusters.

Clustering with k = 1, 2,.., 15
```{r}
wss <- 0
min_n_clusters <- 5
max_n_clusters <- 10
min_n_principal_components <- 1
max_n_principal_components <- 3

km.out <- kmeans(x = pr.out$x[, 1:2], centers = 1, nstart = 20, algorithm = "Lloyd", iter.max = 200)
sil <- silhouette(x = km.out$cluster, dist = dist(pr.out$x[, 1:2]))
plot(sil, col=1, border = NA)

for(Kn in min_n_clusters:max_n_clusters){
  for(Pn in min_n_principal_components:max_n_principal_components){
      km.out <- kmeans(x = pr.out$x[, 1:Pn], centers = Kn, nstart = 20, algorithm = "Lloyd", iter.max = 200)
      sil <- silhouette(x = km.out$cluster, dist = dist(pr.out$x[, 1:Pn]))
      plot(sil, col = 1:Kn, border = NA, main = paste0(Kn, " clusters. ", Pn, " principal components"))
  }
}
```



